---
title: "NYPD Shooting Incident"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Deccription

Shooting incident data in New York Since 2006.

### Source

<https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD>


## Add Library

and library which needed

```{r add_lib, message=FALSE}
library(tidyverse)
```

## Import Data

Import data from website

```{r import_data}
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
```

## Read Data

Read data from csv

```{r read_data}
data <- read.csv(url)
```

## Clean Up Data

Select only focus columns and formatting date

```{r clean_data}
data <- data %>%
  select(OCCUR_DATE, BORO, STATISTICAL_MURDER_FLAG) %>%
  mutate(date = as.Date(OCCUR_DATE, "%m/%d/%Y")) %>%
  select(-c(OCCUR_DATE))
```

Show summary to check if there is missing data.

```{r show_summary}
summary(data)
```

If there is missing BORO data label as unknown

```{r missing_data}
data <- data %>% mutate(BORO = ifelse(BORO != "", BORO, "unknown"))
```


## Group Data By Area

Group data by date and area. Then, count number of case and murder case

```{r transform_data}
area <- data %>% mutate(cases = 1,
                        murder_cases = ifelse(STATISTICAL_MURDER_FLAG == "true", 1, 0)) %>%
  group_by(date, BORO) %>%
  summarize(cases = sum(cases),
            murder_cases = sum(murder_cases))
```

## Transform Data

prepare data for visualization

### Total NY Case

```{r NY_case}
NY_total <- area %>% group_by(date) %>%
  summarize(cases = sum(cases),
            murder_cases = sum(murder_cases)) %>%
  mutate(cases = cumsum(cases),
         murder_cases = cumsum(murder_cases)) %>%
  mutate(murder_percent = murder_cases/cases*100)
```

### Queens Case

```{r queens_case}
queens_case <- area %>% filter(BORO == 'QUEENS') %>%
  select(-c(BORO)) %>%
  ungroup %>%
  mutate(cases = cumsum(cases),
         murder_cases = cumsum(murder_cases)) %>%
  mutate(murder_percent = murder_cases/cases*100)
```

## visualizations

New York cases

```{r NY_plot, echo=FALSE}
NY_total %>%
  ggplot(aes(x = date, y = murder_percent)) +
  geom_line(aes(color = "murder_percent")) +
  geom_point(aes(color = "murder_percent")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "NY murder case ratio", y = NULL)
```

New York cases since 2007

```{r NY_2007_plot, echo=FALSE}
NY_total %>% filter(date > "2007-01-01") %>%
  ggplot(aes(x = date, y = murder_percent)) +
  geom_line(aes(color = "murder_percent")) +
  geom_point(aes(color = "murder_percent")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "NY murder case ratio", y = NULL)
```

cases in Queens since 2007

```{r queens_plot, echo=FALSE}
queens_case %>% filter(date > "2007-01-01") %>%
  ggplot(aes(x = date, y = murder_percent)) +
  geom_line(aes(color = "murder_percent")) +
  geom_point(aes(color = "murder_percent")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "Queens murder case ratio", y = NULL)
```

## Modeling

Create model for predict murder cases by cases


```{r model}
mod <- lm(murder_cases ~ cases, data = NY_total)
```

Predict murder cases with model

```{r predict}
NY_total <- NY_total %>% mutate(predict = predict(mod))
```

Plot predict result

```{r predict_plot, echo=FALSE}
NY_total %>% ggplot() +
  geom_point(aes(x= cases, y = murder_cases, color = "actual")) +
  geom_point(aes(x= cases, y = predict, color = "predicted")) +
  theme(legend.position = "bottom") +
  labs(title = "murder cases from all cases", y = NULL)
```


## Analysis

As you can see from graph 2 (NY cases since 2007) murder rate decrease until 2015. Since 2015 to 2020 murder rate is stable around 19%. But, if you look at graph 3 (Queens cases since 2007) rate is decrease from 2007 then stable at about 18.3% for a while. Until 2010 murder rate is increase to 19-20%.

From this information, I thinks there was something happened in Queens around 2010 which cause murder rate in Queens not keep decreasing until 2015 as same as overall New York City.


## conclusion

New York murder rate overall decrease compare to 2007. But in Queens is difference.

### Bias

Bias of this analysis might be how I calculate murder percentage. Because in very early date there are only few case to calculate compare to latest date. As you can see there is 50% murder rate in graph 1 (NY cases since 2006). So, I prevent this by using data since 2007 which I assumes 2007 have enough cases.

Also, there's personal bias That race dose not effect murder rate at all. I just looked back what is impression with the data and I found that I remove columns about race out because my bias. actually it might be a variable which effect murder rate.

